Supplementary materials to: Describing vocalizations in young children: A big data approach through citizen science annotations

### 2021-03-16 202010_jslhr/paper.Rmd
increased reproducibility

read.csv("../data_analyses/output/key_info.csv")->x
rownames(x)<-x$X

read.csv("../data_analyses/output/chunks_maj_judgments.csv")->chunks

read.csv("../data_analyses/output/zoo_lab_maj_judgments.csv")->data_all

>>>>

read.csv("key_info.csv")->x
rownames(x)<-x$X

read.csv(".chunks_maj_judgments.csv")->chunks

read.csv("zoo_lab_maj_judgments.csv")->data_all

### 2021-03-16 202010_jslhr/sup_mat.Rmd
increased reproducibility

#dictionary that relates chunk to segment created by Chiara
dict <- fromJSON(file="../data_analyses/files_from_elsewhere/dict_4.json") 
dict.simple=unlist(dict)
length(dict.simple) 
dict=data.frame(cbind(dict.simple,gsub(".$","",as.character(names(dict.simple)))))
colnames(dict)<-c("chunk","segmentId_DB")
length(levels(factor(dict$chunk)))
#33728 chunks
length(levels(factor(dict$segmentId_DB)))
#12170 segments
#NOTE!!! THERE ARE TWO CHUNKS MISSING FROM THIS DICTIONARY!!!

#laboratory annotations
read.csv("../data_analyses/files_from_elsewhere/result_final_lisa.csv")->lab_jud
dim(lab_jud)
#11982 segments

@@

#chunk info
read.csv("../data_analyses/output/chunks_maj_judgments.csv",colClasses="factor")->maj_jud
maj_jud$AudioData=gsub(".mp3","",maj_jud$filename)

@@

#but the mytype vector will include n's of each type
mytype <- rep("",n=length(levels(factor(maj_jud$segmentId_DB))))
mytype<-ifelse(mytab[,"Canonical"]>0,paste(mytype,"Ca",mytab[,"Canonical"],sep=""),mytype)
mytype<-ifelse(mytab[,"Non-Canonical"]>0,paste(mytype,"N",mytab[,"Non-Canonical"],sep=""),mytype)
mytype<-ifelse(mytab[,"Crying"]>0,paste(mytype,"Cr",mytab[,"Crying"],sep=""),mytype)
mytype<-ifelse(mytab[,"Laughing"]>0,paste(mytype,"L",mytab[,"Laughing"],sep=""),mytype)
mytype<-ifelse(mytab[,"Junk"]>0,paste(mytype,"J",mytab[,"Junk"],sep=""),mytype)

levels(factor(mytype))
#there are over 350 combinations!! So we won't do that

@@

@@

table(data_all$lab)
table(data_all$Zoon_classif)
mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

@@

```{r prec}
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

```{r rec}
rowsums=rowSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:18) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
rowSums(my_conf_tab)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

@@

@@

```{r}
data_as_td<-left_join(data_all,demo_data,on="ChildID")
# CM with just AS kids
data_AS<-subset(data_as_td, Diagnosis=="AngelmanSyndrome")
mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```

@@

```{r}
# CM with just TD kids
data_TD<-subset(data_as_td, Diagnosis=="Low-RiskControl")
mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```

>>>>

#dictionary that relates chunk to segment created by Chiara
dict <- fromJSON(file="dict_4.json") 
dict.simple=unlist(dict)
length(dict.simple) 
dict=data.frame(cbind(dict.simple,gsub(".$","",as.character(names(dict.simple)))))
colnames(dict)<-c("chunk","segmentId_DB")
length(levels(factor(dict$chunk)))
#33728 chunks
length(levels(factor(dict$segmentId_DB)))
#12170 segments
#NOTE!!! THERE ARE TWO CHUNKS MISSING FROM THIS DICTIONARY!!!

#laboratory annotations
read.csv("lab_annotations.csv")->lab_jud
dim(lab_jud)
#11982 segments

@@

#chunk info
read.csv("chunks_maj_judgments.csv",colClasses="factor")->maj_jud
maj_jud$AudioData=gsub(".mp3","",maj_jud$filename)

@@

# #but the mytype vector will include n's of each type
# mytype <- rep("",n=length(levels(factor(maj_jud$segmentId_DB))))
# mytype<-ifelse(mytab[,"Canonical"]>0,paste(mytype,"Ca",mytab[,"Canonical"],sep=""),mytype)
# mytype<-ifelse(mytab[,"Non-Canonical"]>0,paste(mytype,"N",mytab[,"Non-Canonical"],sep=""),mytype)
# mytype<-ifelse(mytab[,"Crying"]>0,paste(mytype,"Cr",mytab[,"Crying"],sep=""),mytype)
# mytype<-ifelse(mytab[,"Laughing"]>0,paste(mytype,"L",mytab[,"Laughing"],sep=""),mytype)
# mytype<-ifelse(mytab[,"Junk"]>0,paste(mytype,"J",mytab[,"Junk"],sep=""),mytype)
# 
# levels(factor(mytype))
# #there are over 350 combinations!! So we won't do that

@@

# create columns with names that match the following chunks
all_jud$Zoon_classif = all_jud$Answer
all_jud$lab = all_jud$Major_Choice

# for the lab case, simplify
all_jud$lab[all_jud$lab=="Canonical syllables"]<-"Ca"
all_jud$lab[all_jud$lab=="Words"]<-"Ca"
all_jud$lab[all_jud$lab=="Crying"]<-"Cr"
all_jud$lab[all_jud$lab=="Laughing"]<-"L"
all_jud$lab[all_jud$lab=="Don't mark"]<-"J"
all_jud$lab[all_jud$lab=="None"]<-""
all_jud$lab[all_jud$lab=="Non-canonical syllables"]<-"N"

table(all_jud$lab)
table(all_jud$Zoon_classif)

#remove classes with fewer than 10 instances
table(all_jud$Zoon_classif)[table(all_jud$Zoon_classif)<10]
all_jud=all_jud[!(all_jud$Zoon_classif %in% names(table(all_jud$Zoon_classif)[table(all_jud$Zoon_classif)<10])),]

#remove classes with no majority judgment
all_jud=all_jud[all_jud$Zoon_classif !="",]
all_jud=all_jud[all_jud$lab !="",]
dim(all_jud)

all_jud$Zoon_classif=factor(all_jud$Zoon_classif)
all_jud$lab=factor(all_jud$lab, levels=levels(all_jud$Zoon_classif))
```

@@

mycf=confusionMatrix(all_jud$lab, all_jud$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
mycf
```

@@

```{r prec}
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(my_conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
colSums(my_conf_tab)
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","pr")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(pr)), colour = "white") +
  geom_text(aes(label = paste(round(pr),"%")), vjust = -1,size=2) +
#  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Precision")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

```{r rec}
rowsums=rowSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(my_conf_tab)[1]) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
rowSums(my_conf_tab)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall
stall$id=paste(stall$Lab,stall$Zooniverse)
stall=merge(stall,prop_cat[c("id","rec")])
ggplot(data = stall, mapping = aes(y = Lab, x=Zooniverse)) +
 geom_tile(aes(fill= rescale(rec)), colour = "white") +
  geom_text(aes(label = paste(round(rec),"%")), vjust = -1,size=2) +
  geom_text(aes(label = Freq), vjust = 1,size=1) +
  scale_fill_gradient(low = "white", high = "red", name = "Percentage") +
     theme(legend.position = "none") +
  xlab("Zooniverse") + ylab("Lab") +
  ggtitle("Recall")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

## Read in & clean up final data

@@

```{r clean up, echo=F}
#dim(data_all) #11980    16
# remove non-majority labels from lab
data_all[data_all$Num_Agreement>0,]->data_all
#dim(data_all) # 11647    16

# remove non-majority labels from zooniverse
data_all[data_all$Zoon_classif %in% label_options,]->data_all
#dim(data_all) # 11268    16

#and reset the factors for cleanliness
data_all$Zoon_classif=factor(data_all$Zoon_classif, levels=label_options)
data_all$lab=factor(data_all$lab, levels=label_options)
sample_data<-cbind(data_all$lab,data_all$Zoon_classif)

table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100,2)
table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100,2)
```

```{r prepare data by child, echo=F}

#get the ns by child, then calculate the linguistic proportion & canonical proportion, separately for zooniverse & lab
ztab=table(data_all$filename,data_all$Zoon_classif)
z_lp =rowSums(ztab[,c("Canonical","Non-Canonical")])/rowSums(ztab[,-which(colnames(ztab) %in% c("Junk"))])
z_cp =ztab[,c("Canonical")]/rowSums(ztab[,c("Canonical","Non-Canonical")])

ltab=table(data_all$filename,data_all$lab)
l_lp =rowSums(ltab[,c("Canonical","Non-Canonical")])/rowSums(ltab[,-which(colnames(ztab) %in% c("Junk"))])
l_cp =ltab[,c("Canonical")]/rowSums(ltab[,c("Canonical","Non-Canonical")])

#and also get Junk from zooniverse
z_Junk=ztab[,c("Junk")]/rowSums(ztab)

#put all the proportions together
if(sum(rownames(ztab)==rownames(ltab))==dim(ztab)[1]) proportions=cbind(rownames(ztab),z_lp,z_cp,l_lp,l_cp,z_Junk) else print("oops this code needs to be more complex because we don't have the same kids for the two proportions")
colnames(proportions)[1]<-"filename"

#ages=aggregate(data_all$Age,by=list(data_all$ChildID),mean) #this is a weird way of adding ages, since all of the ages for a given child should be the same if there is only one recording, and if there are multiple recordings, then we should not get the mean
#improvement: now we merge with a demo data tab, but note this is merged with child id, so the problem of multiple recs per child is still there

# Created demo_data with filenames. Use filenames instead of childIDs to merge proportions and demo data.

merge(proportions,demo_data_fn,by="filename")->proportions
colnames(proportions)[dim(proportions)[2]]<-"Age"

#cbinding results in text, so we numerize the proportions
for(thisvar in c("z_lp","z_cp","l_lp","l_cp","z_Junk")) proportions[,thisvar]=as.numeric(as.character(proportions[,thisvar]))

#add median Junk
proportions$median_Junk<-ifelse(proportions$z_Junk<median(proportions$z_Junk),"lower_Junk","higher_Junk")

#summary(proportions)

merge(data_all,proportions,all=T)->data_all
```

@@

```{r}
# CM with just AS kids
data_AS<-subset(data_all, Diagnosis=="AngelmanSyndrome")
mycf=confusionMatrix(data_AS$lab, data_AS$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```

@@

```{r}
# CM with just TD kids
data_TD<-subset(data_all, Diagnosis=="Low-RiskControl")
mycf=confusionMatrix(data_TD$lab, data_TD$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
mycf
```
